---
title: "Conduct preliminary analyses"
author: "Christina Bergmann & Alejandrina Cristia"
date: "June 2, 2015"
output:
  md_document:
    variant: markdown_github
---

## Step 1: Load necessary packages

We use two libraries to conduct our meta-analyses: _meta_ and _metafor_. Make sure both are installed and load them with the `library` call in *R*.

```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(meta)
library(metafor)

#Read in raw datafile
library(RCurl)
linkdb = "https://raw.githubusercontent.com/christinabergmann/InWordDB/master/InWordDBnoOutnoRM_ESg.csv"
db.noOutnoRM <- getURL(linkdb)
db.noOutnoRM <- read.csv(textConnection(db.noOutnoRM))
```


## Step 2: General random effects model


This is the first step in a meta-analysis, calculating the general random effects model with no moderators while trimming out the outliers and taking a single effect size per infant group (no repeated measures).

```{r, warning=FALSE, tidy=TRUE}
grlmodel <-rma(ESg, sei=ESg.SE, data=db.noOutnoRM, weighted=TRUE)
summary(grlmodel)
```

The estimate significantly above zero, and in the .2 region, a small but significant effect according to Cohen's (1988) criteria. Furtermore, heterogenity is significant, green light to look at moderators in targeted analyses. 


## Step 3: Inspect funnel plot for asymmetries

```{r}
funnel(grlmodel,xlab="Effect size Hedges' g",pch = 20)
```

